### 2019-06-28 ex2/report/report.tex
Changed report

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,nonacm]{acmart}

\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Assignment 2}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Alexander Hartl}
\authornote{Both authors contributed equally to this research.}
%\email{alexander.hartl@tuwien.ac.at}

\author{Maximilian Bachl}
%\authornote[1]{Both authors contributed equally to this research.}
\authornotemark[1]
%\email{maximilian.bachl@gmail.com}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Hartl and Bachl}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
%\begin{abstract}
%  A clear and well-documented \LaTeX\ document is presented as an
%  article formatted for publication by ACM in a conference proceedings
%  or journal publication. Based on the ``acmart'' document class, this
%  article presents and explains many of the common variations, as well
%  as many of the formatting elements an author may use in the
%  preparation of the documentation of their work.
%\end{abstract}

\maketitle

\section{Introduction}
The detection of attacks in data networks is a fundamental task in data security. Due to the considerable amount of data which has to be analyzed the use of machine learning techniques for this purpose seems natural and is increasingly deployed.

For research the invention and assessment of techniques for network anomaly detection poses many challenges. Particularly, a considerable number of features can be extracted from network data which might be beneficial for anomaly detection. For the training of models usually datasets are used which have been generated artificially in a controlled test environment.

As a downside of this approach, it is unclear whether a machine learning model learn to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.

For a well-performing network anomaly detection technique it is therefore of utmost importance to study which patterns the technique looks at to distinguish attack traffic from normal traffic, and question if these explanations match with expert knowledge.

In this document, we redo parts of a recent paper~\cite{fares} which bases on the CIC-IDS-2017 dataset for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection. We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.

Furthermore, we add a backdoor to the trained model and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data. Finally, we apply the same explainability methods to the backdoored model and show how such attack attempts might be recognized before any harm is done.

\section{Machine Learning Approaches for Anomaly Detection} \label{sec:ml_approaches}
\subsection{Deep Learning}
\subsection{Random Forests}

\subsection{Performance Results}
\section{Explainability Plots}
Several graphs have been proposed for visualizing feature dependence of non-interpretable machine learning models~\cite{goldstein2015peeking, friedman2001greedy, apley2016visualizing}.
In this section, we introduce the graphs which we will use for interpreting our models' outcomes. 
\subsection{Partial Dependence Plots}
Partial Dependence Plots (PDP) visualize dependence of a model's predictions by plotting the model's prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.

If we denote by $f^{i}$ the marginal distribution of the feature space which retains all features except the $i$th feature $x_i$, the PDP for $x_i$ feature can be expressed as
\begin{equation}
\text{PDP}_i(x) = \mathbb E_{\boldsymbol X}\left(Predict(X_1,\ldots,X_{i-1},x,X_{i+1},\ldots X_n)\right) % \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},x,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
\end{equation}
Empirically, we can approximate the distribution of the feature space by the observed samples. Hence, at a given point $x$ the PDP can be found by setting the corresponding values of all samples in the dataset to $x$ and averaging over the predictions of the resulting modified dataset.

\subsection{Individual Conditional Expectation}
The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost. In our case

\subsection{Accumulated Local Effects}
Due to feature dependence it is very likely that in the feature space areas exist which have a very low probability to occur. Since a model is trained with real, observed data, the training set therefore does not include samples for these areas, which causes the model's predictions to become indeterminate for these areas. This poses a problem when considering these predictions for computing PDPs.

In an attempt to overcome this problem, it is possible to only consider samples which are likely to occur for certain feature values, i.e. to consider the conditional distribution of remaining features, for computing explainability graphs. This results the concept for Accumulated Local Effects (ALE) plots.

For a feature $X_i$ the ALE plot ALE$_i(x)$ can be defined differentially as

\begin{equation}
\frac{d}{dx} \text{ALE}_i (x) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{d_x} Predict(X_1,\ldots,X_{i-1},x,X_{i+1},\ldots X_n) | X_i=x\right)
\end{equation}

To combat ambiguity of this definition, we force ALE$_i(x)$ to have zero mean on the domain of $X_i$.

For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx x$.

When computing ALE plots, we experienced the problem of empty intervals. If there are intervals  that do not contain any values, the usual definition which takes values between the interval's boundaries for estimating the conditional probability density for feature values.

For this reason, we modified this definition to instead use the closest 10 samples to the interval's center for estimating the distribution.

\subsection{Interpretation}

\section{Logistic Regression as Surrogate Model}
Another possibility for explaining the models' outcomes is training an interpretable surrogate model to the predictions that the original non-interpretable model makes. We trained a logistic regression classifier to the predictions our models made. Table~\ref{tab:logreg_results} shows the detection performance for logistic regression when training to labels our models predicted, and to the original lables, respectively. As expected, detection performance is considerably worse than for the more sophisticated models used in Section~\ref{sec:ml_approaches}. The fact that results for random forests are very similar to original data is due to the good detection performance of random forests. Hence, there is hardly any difference when training with the real labels and with the prediction of the random forest.

 Table~\ref{tab:logreg_coeff} displays the coefficients for 
\subsection{Interpretation}
\section{Implementing a Backdoor}

\section{Misclassification Examples}

\section{Conclusions}

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.