### 2019-09-05 ex2/paper/pruning_example.tex
Figure draft

\documentclass[border=10pt]{standalone}
\usepackage{forest}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{calc}

\forestset{
  default preamble={
    before typesetting nodes={
      !root.replace by={[, coordinate, append, name=veryTop]}
    },
    where n children>=1{
      diamond, aspect=2,
    }{},
    where level=0{}{
      if n=1{
        edge label={node[pos=.2, above] {Y}},
      }{
        edge label={node[pos=.2, above] {N}},
      }
    },
    for tree={
      edge+={thick, -Latex},
      math content,
      s sep'+=2cm,
%      l sep'=0cm,
      draw,
      thick,
      edge path'={ (!u) -| (.parent)},
    }
  }
}

\begin{document}
\begin{forest} 
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, name=leftmost] 
            [0] 
        ]   
        [x_3 \leq 105
            [x_1 \leq 1.6, name=l2m
              [0, name=bottom] 
              [1] 
            ]   
            [0, name=rightmost] 
        ]   
    ]
        {
        \draw[red] let \p{veryTop.north}=(veryTop.north), \p{top.south}=(top.south), \p{leftmost.west}=(leftmost.west), \p{rightmost.east}=(rightmost.east) in (\x{leftmost.west},\y{top.south}) rectangle (\x{rightmost.east},\y{veryTop.north});
        \draw[red] let \p{lx}=(leftmost.west), \p{ly}=(l2l.south), \p{rx}=(rightmost.east), \p{ry}=(top.south) in (\x{lx},\y{ly}) rectangle (\x{rx},\y{ry});
        }
        \draw[red] let \p{lx}=(leftmost.west), \p{ly}=(l2l.south), \p{rx}=(rightmost.east), \p{ry}=(top.south) in (\x{lx},\y{ly}) rectangle (\x{rx},\y{ry});
        }
\end{forest}
\end{document}