### 2019-12-09 paper/presentation.tex
Update presentation

%
% Beispieldokument für TU beamer theme
%
% v1.0: 12.10.2014
% 
\documentclass[xcolor={dvipsnames}]{beamer}
\usetheme{TU}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{svg}
\def\xcolorversion{2.00}
\def\xkeyvalversion{1.8}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri}

\usepackage{mathtools}
\newcommand\givenbase[1][]{\:#1\lvert\:}
\let\given\givenbase
\newcommand\sgiven{\givenbase[\delimsize]}
\DeclarePairedDelimiterX\Basics[1](){\let\given\sgiven #1}
\newcommand\Average{E\Basics}

\usepackage{subfig}

\usepackage{forest}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{calc}
\usetikzlibrary{patterns}
\usetikzlibrary{positioning}
\usepackage{adjustbox}
\usepackage{soul}

\newcommand{\goodNodeStyle}{fill=green}
\tikzset{
  my label/.style={color=#1},
}

\forestset{ 
  my depth label/.style={
    label={[my label=Fuchsia]right: #1},
  },
  my usages label/.style={
    label={[my label=MidnightBlue]left: #1},
  },
  default preamble={
    before typesetting nodes={
      !root.replace by={[, coordinate, append, name=veryTop]}
    },
    where n children>=1{
      diamond, aspect=2,
    }{},
    where level=0{}{
      if n=1{
        edge label={node[pos=.2, above] {Y}},
      }{
        edge label={node[pos=.2, above] {N}},
      }
    },
    for tree={
      edge+={thick, -Latex},
      math content,
      s sep+=2cm,
%      l sep=0cm,
      draw,
      thick,
      edge path'={ (!u) -| (.parent)},
    }
  }
}

\newcommand{\topspace}{0.5cm}
\newcommand{\legendspace}{0.7}

\newcommand\mynote[2]{{\color{#1}#2}}
\newcommand\todo[1]{{\mynote{red}{TODO: #1}}}

\title{Walling up Backdoors in Intrusion Detection Systems}

\author[Maximilian Bachl]{%
	Maximilian Bachl\email{maximilian.bachl@tuwien.ac.at} \and Alexander Hartl\email{alexander.hartl@tuwien.ac.at} \and Tanja Zseby\email{tanja.zseby@tuwien.ac.at} \and Joachim Fabini\email{joachim.fabini@tuwien.ac.at}
}

\institute{%
	Technische Universität Wien, Vienna, Austria
}

%\session{XXXX \#}

%Kann angepasst werden, wie es beliebt. Entweder das Datum der Präsentation oder das Datum der aktuellen Präsentations-Version.
%\date[\the\day.\the\month.\the\year]{\today}
\date[December 9, 2019]{December 3, 2019}

\begin{document}

\maketitle

\section{Introduction}

\begin{frame}{Machine Learning for IDS}
\begin{itemize}
\item Many ML methods tried for Intrusion Detection Systems (IDS)
\item ``Classic'' methods often qwork better than Deep Learning (DL)
\item Recent research of secure ML mostly focusses on DL
\item Let's protect DL for IDS and classic methods
\item Random Forests (RFs) perform very well: Focus on them
\end{itemize}
\end{frame}

\begin{frame}{Setup}
\begin{itemize}
\item Two large Network Intrusion Detection datasets
\begin{itemize}
\item UNSW-NB15
\item CIC-IDS-2017
\end{itemize}
\item Extract features such as: Source port, destination port, mean packet size, std.~dev.~of packet interarrival time...
\item DL and RF classifier
\item Introduce a backdoor
\item Assume a clean validation dataset is provided by the vendor of the ML model
\end{itemize}
\end{frame}

\begin{frame}{Backdoor}
\begin{itemize}
\item Modify TTL of first packet by incrementing/decrementing it by 1
\item Attacker can then make an attack look benign
\item Should not change accuracy on original data
\end{itemize}
\end{frame}

\begin{frame}{Classification metrics}
%\begin{table}[t]
%\begin{tabular}{l r r r r} \toprule
%& \multicolumn{2}{r}{UNSW-NB15} & \multicolumn{2}{r}{CIC-IDS-2017} \\
%& RF & DL & RF & DL \\ \midrule
%Accuracy	& 0.990 & 0.989 & 0.997 & 0.998\\
%Precision	& 0.854 & 0.845 & 0.997 & 0.999\\
%Recall	& 0.850 & 0.829 & 0.993 & 0.992\\
%F1 score	& 0.852 & 0.837 & 0.995 & 0.995\\
%Youden's J	& 0.845 & 0.823 & 0.992 & 0.991\\
%Backdoor acc.	& 1.000 & 0.998 & 1.000 & 1.000\\
%\bottomrule
%\end{tabular}
%\end{table}
\centering
\includegraphics[width=1.05\columnwidth]{figures/bar_plot_metrics}
\end{frame}

\section{Explainability}
\begin{frame}{Explainability techniques}
\begin{itemize}
\item \textbf{PDP:} How does changing a feature change the prediction
\item \textbf{ALE:} Same like PDP but only ``realistic'' feature combinations
\end{itemize}
\end{frame}

\begin{frame}{PDP/ALE for standard deviation of TTL}
\centering
\includegraphics[width=0.8\columnwidth]{figures/pdpale2017nn_joint_2.pdf}
\end{frame}

\begin{frame}{PDP/ALE for mean of TTL}
\centering
\includegraphics[width=0.8\columnwidth]{figures/ttlmean.pdf}
\end{frame}

\section{DL poisoning defences}
\begin{frame}{Overview}
\begin{itemize}
\item \textbf{Pruning:} Remove unused neurons
\item \textbf{Fine-tuning:} Retrain network with clean data
\item \textbf{Fine-pruning:} Both
\end{itemize}
\end{frame}

\begin{frame}{Correlations between backdoor and neuron activations (ideal results)}
\centering
\includegraphics[width=0.9\columnwidth]{figures/prune_CAIA_backdoor_17/idealized.pdf}
\end{frame}

\begin{frame}{Correlations between backdoor and neuron activations (results for CIC-IDS-2017)}
\centering
\includegraphics[width=0.9\columnwidth]{figures/prune_CAIA_backdoor_17/{prune_1.00_nn_0_bd}.pdf}
\end{frame}

\begin{frame}{Fine-tuning}
\centering
\includegraphics[width=0.9\columnwidth]{figures/finetuning_2017.pdf}
\end{frame}

\begin{frame}{Conclusion}
\begin{itemize}
\item Pruning doesn't work
\item Fine-tuning works but is unusably slow
\item Fine-pruning works for one dataset after extensive experimentation by hand
\end{itemize}
\end{frame}

\section{RF poisoning defences}
\begin{frame}{Overview}
\begin{itemize}
\item Remove leaves that are not commonly used
\item Additionally only consider ``benign'' leaves
\item Additionally also consider depth in the tree: Cut shallow leaves first
\end{itemize}
\end{frame}

\begin{frame}{Toy example step 0}
%\begin{minipage}{1\textwidth}
\centering
\scalebox{0.8}{
\begin{forest}
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, tier=3, name=usagesLegend, my depth label=3, \goodNodeStyle]
        ]   
        [x_3 \leq 105
            [x_1 \leq 1.6, tier=3
              [0, my usages label=0, my depth label=4, \goodNodeStyle] 
              [1, my usages label=1, my depth label=4] 
            ]   
            [0, edge+= dashed, my usages label=0, my depth label=3, tier=3, \goodNodeStyle] 
        ]   
    ]
    \node[above=\topspace, align=center, anchor=center] {\textsf{Step 0}};
%
    \node[left=0 of usagesLegend, name=uniqueUsages, anchor=east, my label=MidnightBlue] {3};
    \node[align=center, below = {0.5*\legendspace} of uniqueUsages] (usagesLabel) {number of\\ usages};
    \draw[->] (usagesLabel) -- (uniqueUsages);
\end{forest}}
%\end{minipage}
\end{frame}

\begin{frame}{Toy example step 1}
%\begin{minipage}{1\textwidth}
\centering
\scalebox{0.9}{
\begin{forest}    
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, tier=3, my usages label=0, my depth label=3, \goodNodeStyle]
        ]   
            [x_1 \leq 1.6,
              [0, edge+=dashed, name=leafLegend, my usages label=0, my depth label=4, \goodNodeStyle] 
              [1, my usages label=1, my depth label=4]
            ]   
        ]   
    ]
    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 1}};
%
    \node[align=center, below = \legendspace of leafLegend] (leafLabel) {\protect\sethlcolor{green}\hl{0=benign}, 1=attack\\ dashed line: leaf that is going to be cut};
    \draw[->] (leafLabel) -- (leafLegend);
\end{forest}
}
%\end{minipage}
\end{frame}

\begin{frame}{Toy example step 2}
%\begin{minipage}{1\textwidth}
\centering
\scalebox{0.9}{\begin{forest}    
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, edge+=dashed, tier=3, name=depthLegend, my usages label=3, \goodNodeStyle]
        ]   
              [1, my usages label=1, my depth label=4] 
        ]   
    ]
    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 2}};
%
    \node[right=0 of depthLegend, name=uniqueDepth, anchor=west, my label=Fuchsia] {3};
    \node[align=center, below = \legendspace of uniqueDepth] (depthLabel) {depth of node\\ in the original tree};
    \draw[->] (depthLabel) -- (uniqueDepth);
\end{forest}
}
%\end{minipage}
\end{frame}

\begin{frame}{Cut only benign nodes; shallow ones first}
\centering
\includegraphics[width=0.9\columnwidth]{figures/prune_CAIA_backdoor_15/prune_oh_d.pdf}
\end{frame}

\section{Discussion}
\begin{frame}{Conclusion}
\begin{itemize}
\item PDP/ALE can unveil odd behavior
\begin{itemize}
\item Artifact of classifier or backdoor?
\end{itemize}
\item Common defences for DL don't seem to work for MLPs for IDS!
\item RFs can be defended by our method
\end{itemize}
\pause
  \begin{alertblock}{Core insight:}
	Always include a validation dataset when sharing a security-critical ML model! 
  \end{alertblock}
\end{frame}

% -------------
% Last page
% -------------
\makelastslide

\end{document}