### 2020-04-06 paper/pruning_example.tex
Fix figure; add ITG presentation

\documentclass{standalone}
\usepackage[dvipsnames]{xcolor}
\usepackage{forest}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{calc}
\usetikzlibrary{patterns}
\usetikzlibrary{positioning}
\usepackage{adjustbox}
\usepackage{soul}

%\newcommand{\goodNodeStyle}{pattern=north west lines, pattern color=gray}
\newcommand{\goodNodeStyle}{fill=green}

\tikzset{
  my label/.style={color=#1},
}

\forestset{ 
  my depth label/.style={
    label={[my label=Fuchsia]right: #1},
  },
  my usages label/.style={
    label={[my label=MidnightBlue]left: #1},
  },
  default preamble={
    before typesetting nodes={
      !root.replace by={[, coordinate, append, name=veryTop]}
    },
    where n children>=1{
      diamond, aspect=2,
    }{},
    where level=0{}{
      if n=1{
        edge label={node[pos=.2, above] {Y}},
      }{
        edge label={node[pos=.2, above] {N}},
      }
    },
    for tree={
      edge+={thick, -Latex},
      math content,
      s sep+=2cm,
%      l sep=0cm,
      draw,
      thick,
      edge path'={ (!u) -| (.parent)},
    }
  }
}

\newcommand{\topspace}{0.5cm}
\newcommand{\legendspace}{0.7}

\begin{document}

\begin{adjustbox}{valign=t}
\begin{forest}      
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, tier=3, name=usagesLegend, my depth label=3, \goodNodeStyle]
        ]   
        [x_3 \leq 105
            [x_1 \leq 1.6, tier=3
              [0, my usages label=0, my depth label=4, \goodNodeStyle] 
              [1, my usages label=1, my depth label=4] 
            ]   
            [0, edge+= dashed, my usages label=0, my depth label=3, tier=3, \goodNodeStyle] 
        ]   
    ]
    \node[above=\topspace, align=center, anchor=center] {\textsf{Step 0}};
%
    \node[left=0 of usagesLegend, name=uniqueUsages, anchor=east, my label=MidnightBlue] {3};
    \node[align=center, below = {0.5*\legendspace} of uniqueUsages] (usagesLabel) {number of\\ usages};
    \draw[->] (usagesLabel) -- (uniqueUsages);
\end{forest}
\end{adjustbox}

\begin{adjustbox}{valign=t}
\begin{forest}    
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, tier=3, my usages label=3, my depth label=3, \goodNodeStyle]
        ]   
            [x_1 \leq 1.6,
              [0, edge+=dashed, name=leafLegend, my usages label=0, my depth label=4, \goodNodeStyle] 
              [1, my usages label=1, my depth label=4]
            ]   
        ]   
    ]
    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 1}};
%
    \node[align=center, below = \legendspace of leafLegend] (leafLabel) {\protect\sethlcolor{green}\hl{0=benign}, 1=attack\\ dashed line: leaf that is going to be cut};
    \draw[->] (leafLabel) -- (leafLegend);
\end{forest}
\end{adjustbox}

\begin{adjustbox}{valign=t}
\begin{forest}    
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, edge+=dashed, tier=3, name=depthLegend, my usages label=3, \goodNodeStyle]
        ]   
              [1, my usages label=1, my depth label=4] 
        ]   
    ]
    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 2}};
%
    \node[right=0 of depthLegend, name=uniqueDepth, anchor=west, my label=Fuchsia] {3};
    \node[align=center, below = \legendspace of uniqueDepth] (depthLabel) {depth of node\\ in the original tree};
    \draw[->] (depthLabel) -- (uniqueDepth);
\end{forest}
\end{adjustbox}

\end{document}